---
title: "Assignment 1"
output: html_document
---

```{r}
setwd("C:/Users/HP/Desktop/statistical data analysis/A1")
```


##Task 2.1

```{r}
install.packages("tidyverse")
library(tidyverse)
ed_data_link <- 'govhack3.csv'

top_row <- read.csv(ed_data_link, col.names = FALSE , nrows = 1,header=FALSE)
top_row<-top_row[top_row[,1]!="",1]

second_row <- read.csv(ed_data_link, nrows = 1)

column_names <- second_row %>% unlist(., use.names=FALSE) %>% as.character()%>%
  make.unique(., sep = "__") # double underscore

column_names[2:8] <- str_c(column_names[2:8], '0', sep='__')

daily_attendance <- read_csv(ed_data_link, skip = 2, col_names = column_names)

sapply(daily_attendance,class) # data types of columns

top_row   # list of hospitals
```
##Task 2.2

Make the data tidy using tidyverse gathering and spreading operations
Fixing variable types
Arranging data in chronological order
```{r}
tidy1 <- daily_attendance %>% 
  gather(key='All',value,Attendance__0:Tri_5__8)%>%separate(col='All', into=c('All','Hospital'),sep='__')%>%spread(key='All',value)%>%arrange(Hospital, Date)

# fixing variable types

tidy2<-tidy1%>%mutate(Hospital = factor(Hospital),
         Admissions =as.integer(Admissions),Attendance=as.integer(Attendance),Tri_1=as.integer(Tri_1),Tri_2=as.integer(Tri_2),Tri_3=as.integer(Tri_3),Tri_4=as.integer(Tri_4),Tri_5=as.integer(Tri_5),Date = as.Date(Date, "%d-%B-%Y"))

#Arranging data in chronological order
tidy_data<-tidy2%>%arrange(Hospital,Date)
```
Analyse the data and know which entries are complete and which entries contain NA values.
```{r}
Tri_1=tidy_data$Tri_1
Tri_2=tidy_data$Tri_2
Tri_3=tidy_data$Tri_3
Tri_4=tidy_data$Tri_4
Tri_5=tidy_data$Tri_5
Admissions=tidy_data$Admissions
Attendance=tidy_data$Attendance
Hospital=tidy_data$Hospital

length(Tri_1[is.na(Tri_1)])   #2600 incomplete entries
length(Tri_2[is.na(Tri_2)])   #371 incomplete entries
length(Tri_3[is.na(Tri_3)])   #85 incomplete entries
length(Tri_4[is.na(Tri_4)])   #0 incomplete entries
length(Tri_5[is.na(Tri_5)])   #448 incomplete entries
length(Admissions[is.na(Admissions)])  #87 incomplete entries
length(Attendance[is.na(Attendance)])  #0 incomplete entries

nrow(tidy_data[complete.cases(tidy_data),]) # 601 complete entries


#checking that in how many entries attendance is equal to sum of all triage from complete entries.

complete<-tidy_data[complete.cases(tidy_data),]

nrow(complete[(complete$Attendance==complete$Tri_1+complete$Tri_2+complete$Tri_3+complete$Tri_4+complete$Tri_5),])  #532 

# from the above result we can safely assume that attendance is the sum of all triage and this can used to impute the NA values.

```

Imputations are done by taking account of above result i.e attendance=sum of all triage, and incorporating the fact that NA values are less than 3 as mentioned by the emergency department in the description of the data.

```{r}
############# imputations #############################33



for(i in 1:nrow(tidy_data)){
  if(is.na(Tri_1[i])&!is.na(Tri_2[i])&!is.na(Tri_3[i])&!is.na(Tri_4[i])&!is.na(Tri_5[i])){
    k<-tidy_data$Attendance[i]-(tidy_data$Tri_2[i]+tidy_data$Tri_3[i]+tidy_data$Tri_4[i]+tidy_data$Tri_5[i])
    if(k<3){tidy_data$Tri_1[i]<-k}
    else{tidy_data$Tri_1[i]<-2}
  }
}

Tri_1=tidy_data$Tri_1
length(Tri_1[is.na(Tri_1)])  # reduced to 728 incomplete entries of Triage_1 from 2600

nrow(tidy_data[complete.cases(tidy_data),]) #2472 complete cases


nrow(tidy_data[!is.na(Tri_1)&!is.na(Tri_2)&!is.na(Tri_3)&!is.na(Tri_4)&is.na(Tri_5),])     # 84 rows contain only missing Triage_5 out of all Triage

for(i in 1:nrow(tidy_data)){
  if(!is.na(Tri_1[i])&!is.na(Tri_2[i])&!is.na(Tri_3[i])&!is.na(Tri_4[i])&is.na(Tri_5[i])){
    k<-tidy_data$Attendance[i]-(tidy_data$Tri_2[i]+tidy_data$Tri_3[i]+tidy_data$Tri_4[i]+tidy_data$Tri_1[i])
    if(k<3){tidy_data$Tri_5[i]<-k}
    else{tidy_data$Tri_5[i]<-2}
  }
}

Tri_5=tidy_data$Tri_5
length(Tri_5[is.na(Tri_5)])  # 364 incomplete entries of Triage_5 from 448

nrow(tidy_data[complete.cases(tidy_data),])  # 2556 complete cases


# replacing the rest of NA values by Zero 

tidy_data$Admissions[is.na(Admissions)]<-0
tidy_data$Tri_1[is.na(Tri_1)]<-0
tidy_data$Tri_2[is.na(Tri_2)]<-0
tidy_data$Tri_3[is.na(Tri_3)]<-0
tidy_data$Tri_4[is.na(Tri_4)]<-0
tidy_data$Tri_5[is.na(Tri_5)]<-0

```

##Task 3.1

```{r}
Tri_1=tidy_data$Tri_1
Tri_2=tidy_data$Tri_2
Tri_3=tidy_data$Tri_3
Tri_4=tidy_data$Tri_4
Tri_5=tidy_data$Tri_5
Admissions=tidy_data$Admissions
Attendance=tidy_data$Attendance
Hospital=tidy_data$Hospital
Date=tidy_data$Date

t<-tidy_data[Hospital==4,]
print(top_row[5])
print(paste0('Total Admissions:', sum(t$Admissions))) #39400 
print(paste0('Total Attendances:',sum(t$Attendance))) #65871

```

##Task3.2
```{r}
t %>% 
  ggplot(aes(x = t$Attendance)) +
  geom_histogram(
  )

```

##Task3.3
```{r}
t%>%mutate(day=weekdays(Date))%>%mutate(day=factor(day, levels= c( "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday","Sunday")))%>%group_by(day)%>%summarise(mean(Attendance))%>%plot(xlab='',ylab='Attendace',las=2, main='Attendance average according to weekday ')

```

##Task3.4
```{r}

m=sapply(t[3:9],mean) # mean of each variable
v=sapply(t[3:9],var) # variance of each variable

m
v
```


##Task 4.1

```{r}
install.packages('fitdistrplus')
library(fitdistrplus)
library(magrittr)

estimate1 <- t %$%
  fitdist(data = Tri_1, 
          distr = "pois")

estimate2 <- t %$%
  fitdist(data = Tri_1, 
          distr = "nbinom")

t %>% 
  ggplot(aes(x = t$Tri_1)) +
  geom_histogram(
  )
estimate1[["loglik"]]
estimate2[["loglik"]]

estimate1 %>% plot
estimate2 %>% plot

```

